Adaptive evolution with aneuploidy and mutation

### 2021-06-27 ms/ms.tex
added figures demonstrating growth curve estimation

@@

\paragraph{Parameter inference.} To infer model parameters, we use approximate Bayesian computation~\citep{Sunnaker2013} with a sequential Monte-Carlo scheme, or ABC-SMC~\citep{Sisson2009}, implemented in the \texttt{pyabc} Python 
package~\citep{Klinger2018}.
Briefly, this approach uses numerical stochastic simulations of the model to infer a posterior distribution over the model parameters. This approach is applicable for likelihood-free, simulation-based inference~\citep{Cranmer2020}, that is, for estimating a posterior distribution when a likelihood function cannot be computed. It is therefore suitable to our case, in which the likelihood function can only be approximated from simulations, and cannot be directly computed. 
% TODO ILIA please add all configuration details, similar to the relevant section in the NPI paper. Discuss priors, convergence tests, etc.

\paragraph{Prior distributions.} The prior distributions of $w_{2n+1}=1-c+b$, $w_{2n+1^*}=(1+s)(1-c)+b$ and $w_{2n^*}=1+s$ were obtained by estimating prior distributions for $1-c$, $1-c+b$ and $1+s$ from growth curves data. These growth curve data were previously obtained by~\citet{Yona2012} using mono-culture growth experiments.
The raw growth curves data were not published before, but were used to produce figures 3C, 4A, and S2 in \citet{Yona2012}.
We used \texttt{Curveball}, a dedicated method for predicting results of competition experiments from growth curve data~\citet{Ram2019}. \texttt{Curveball} takes growth curves of two strains growing separately in mono-culture and predicts how they would grow in a mixed-culture, that is, it predicts the results of a competition assay. From these predictions, relative fitness values can be computed.  
We used growth curves of $2n+1$ and $2n$ in \SI{39}{\celsius} and in \SI{30}{\celsius} to estimate $1-c+b$ and $1-c$, respectively. The latter assumes that the cost of aneuploidy is the same in \SI{39}{\celsius} and \SI{30}{\celsius}; this might not be correct, but we only assume this in order to generate a prior distribution for the fitness values. We also used growth curves of $2n+1$ and $2n^*$ in \SI{39}{\celsius} in \SI{39}{\celsius} to estimate $1+s$. 
Because \texttt{Curveball} uses a maximum-likelihood approach to estimate model parameters, we were able to estimate a distribution of relative fitness values by sampling from a truncated multivariate normal distribution defined by the maximum-likelihood covariance matrix. Thus, we sampled 10,000 values for $1-c+b$, $1-c$, and $1+s$, which we used as prior distributions. Note that the first and last values correspond to $w_{2n+1}$ and $w_{2n}$, whereas $w_{2n+1^*}=(1-c)w_{2n^*}+w_{2n+1}-(1-c)$.
%TODO ILIA please report here the point estimate and 95% CI for the three distributions, I only have for 1-c+b: 0.9863 (95% CI 0.9790 - 0.9939).
%TODO ILIA describe the priors for the mutation and aneuploidy rate.

@@

>>>>

% Supplementary
\newcommand{\beginsupplement}{%
      	\setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
}

@@

\paragraph{Parameter inference.} To infer model parameters, we use approximate Bayesian computation~\citep{Sunnaker2013} with a sequential Monte-Carlo scheme, or ABC-SMC~\citep{Sisson2009}, implemented in the \texttt{pyABC}\footnote{https://pyabc.readthedocs.io} Python 
package~\citep{Klinger2018}.
Briefly, this approach uses numerical stochastic simulations of the model to infer a posterior distribution over the model parameters. It is a method of likelihood-free, simulation-based inference~\citep{Cranmer2020}, that is, for estimating a posterior distribution when a likelihood function cannot be computed. It is therefore suitable to our case, in which the likelihood function can only be approximated from simulations, and cannot be directly computed. 
% TODO ILIA please add all configuration details, similar to the relevant section in the NPI paper. Discuss priors, convergence tests, etc.

\paragraph{Prior distributions.} The prior distributions of $w_{2n+1}=1-c+b$, $w_{2n+1^*}=(1+s)(1-c)+b$ and $w_{2n^*}=1+s$ were obtained by estimating $1-c$, $1-c+b$ and $1+s$ from growth curves data. These growth curve data were previously obtained by~\citet{Yona2012} using mono-culture growth experiments.
The raw growth curves data were not published before, but were used to produce figures 3C, 4A, and S2 in \citet{Yona2012}.
We used \texttt{Curveball}\footnote{https://curveball.yoavram.com}, a dedicated method for predicting results of competition experiments from growth curve data~\citep{Ram2019}. \texttt{Curveball} takes growth curves of two strains growing separately in mono-culture and predicts how they would grow in a mixed culture, that is, it predicts the results of a competition assay.
From these predictions, relative fitness values can be computed.  
We used growth curves of $2n+1$ and $2n$ in \SI{39}{\celsius} and in \SI{30}{\celsius} to estimate $1-c+b$ and $1-c$, respectively. The latter assumes that the cost of aneuploidy is the same in \SI{39}{\celsius} and \SI{30}{\celsius}; this might be incorrect, but we only assume this to generate a prior distribution for the fitness values. We also used growth curves of $2n+1$ and $2n^*$ in \SI{39}{\celsius} to estimate $1+s$. 
Because \texttt{Curveball} uses a maximum-likelihood approach to estimate model parameters, we were able to estimate a distribution of relative fitness values by sampling from a truncated multivariate normal distribution defined by the maximum-likelihood covariance matrix. Thus, we sampled 10,000 values for $1-c+b$, $1-c$, and $1+s$, which we used as prior distributions. Note that the first and last values correspond to $w_{2n+1}$ and $w_{2n}$, whereas $w_{2n+1^*}=(1-c)(w_{2n^*}-1)+w_{2n+1}$.
See Figures~\ref{fig:growth-curves-30deg} and~\ref{fig:growth-curves-39deg}.
%TODO ILIA please report here the point estimate and 95% CI for the three distributions, I only have for 1-c+b: 0.9863 (95% CI 0.9790 - 0.9939).
%TODO ILIA describe the priors for the mutation and aneuploidy rate.

@@

\section*{Supplementary Material}
\beginsupplement % https://support.authorea.com/en-us/article/how-to-create-an-appendix-section-or-supplementary-information-1g25i5a/

%%%% Fig - Curveball 30deg %%%
% generated with growth_curves.ipynb
\begin{figure}[h]
    \centering
	\includegraphics[width=0.5\textwidth]{../figures/evo39_fitness_30deg.pdf}
    \caption{
    \textbf{Fitness estimation from \SI{30}{\celsius}.}
    } 
    \label{fig:growth-curves-30deg}
\end{figure}

%%%% Fig - Curveball 39deg %%%
% generated with growth_curves.ipynb
\begin{figure}[h]
    \centering
	\includegraphics[width=0.5\textwidth]{../figures/evo39_fitness_39deg.pdf}
    \caption{
    \textbf{Fitness estimation from \SI{39}{\celsius}.}
    } 
    \label{fig:growth-curves-39deg}
\end{figure}
